\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}

### 2018-03-29 redaction/dhp_july2017.tex
changement texte

\begin{abstract}
This paper investigates the underlying sources behind the downward trend of international transport costs. In the same spirit as \cite{hummels2007}, we identify the respective roles of the reduction in transport costs ``per se'' and changes in trade composition effects. This drives us to provide a careful modeling of international transport costs, that allows for both ad-valorem (multiplicative) and per-unit (additive) components, exploiting information contained in the US imports flows database over 1974-2013 distinguished by transport mode, air and vessel. In contrast to \cite{hummels2007}, trade composition effects (inside each transport mode) do not matter much and tend to amplify (rather than reduce) the decrease in pure transport costs. This suggests that transport costs ``per se'' declined even less than argued by \cite{hummels2007}. Importantly, this difference of results can be attributed to the new way of modeling the additive component of transport costs we offer. Further, in quantitative terms, we find that the additive component of transport costs is sizeable, representing between 40 and 50\% of overall transport costs. In both aspects, our results point the importance of the additive component in accounting for international transport costs.
\end{abstract}

@@

This focus is shared with part of the international trade literature that studies the patterns of transport costs over time, such as \cite{hummels2007} and \cite{Behar_Venables} (among others). Many argue that transport costs have substantially decreased with technological advance in transportation (see \citealp{levinson2016box}), infrastructure development and new communication technologies (see \citealp{Lafourcade_Thisse}). %According to \cite{Glaeser04}, over the twentieth century, the cost of moving goods inland have declined by over 90\% in real terms.
Yet, distance is still a very important determinant of trade flows and some have noticed that progress in transportation techniques must be compared with technical progress in production. The constant-dollar cost of transport has declined, but so has the constant-dollar cost of production. Furthermore, the quality of transport services have increased. As a result, transport costs as a share of export prices have declined less quickly than expected  (\citealp{hummels2007}).\smallskip

According to \cite{hummels2007}, this lower-than-expected decline of transport costs is partly explained by a composition effect. Based on US imports over 1974-2004, he finds that transport costs (i.e., \textit{ceteris paribus} for given goods, mode, and trade routes) have decreased in both air and maritime shipment, but less so than ``per se'' transport costs because trade composition effects have partly offset this ``per se'' transport costs decrease, thereby attenuating the overall transport costs downward trend. In this paper, we investigate the robustness of this result. As a major contribution, and in contrast to \cite{hummels2007}, we find that trade composition effects play a minor role in accounting for the trend patterns of transport costs. Rather, the evolution of transport costs ``per se'' constitutes the main driving force behind the overall transport costs decrease, in particular for air transport. In maritime shipment, trade composition effects matter more, but they amplify (rather than offset) the reduction in pure transport costs, suggesting that transport costs ``per se'' declined less than argued by {hummels2007}.\medskip

To prove this point we provide a careful modelling of transport costs. Following \citet{samuelson1954}, standard models of international trade have usually modeled trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis).\footnote{Strictly speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded. As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as it is commonly done in the literature.} However, a recent strand of the literature points out to the existence of an additive component, that is, a cost per unit traded (see, among others, \citealp{Irrazabal_2015} or \citealp{martin2012}). Based on US sectoral data, our own estimates also bring out the importance of additive costs in international transport costs. This is obtained by providing an empirical decomposition of the structure of transport costs over time, in which we explicitly distinguish between ad-valorem and additive costs. To do so, we exploit information contained in the US imports flows over 1974-2013, from which we get a measure of international transport costs as the difference between the import and the export prices, on a year-product-partner country basis. This database is the same as \cite{hummels2007}, extended over more recent years (2005-2013). \smallskip

@@

Second, we find that the main source of the downward trend in overall transport costs (that decreased by roughly 50\% in air shipping and 60\% in vessel shipping over 1974-2013) does come from the reduction of transport cost ``per se''. Trade composition effects only have a limited influence, and when they matter (in maritime transport), they amplify the reduction of ``pure'' transport costs, suggesting that ``per se'' trade costs declined less than observed trade costs. This stands in contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004). Importantly, we show that the difference of results can be attributed to the new way of modeling of the additive component we offer. Allowing the share of additive costs to vary  across time, products and country partners (rather than being constant as in \citealp{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period. Our result thus deepens the paradox of the less-than-expected decline of transport costs.\smallskip

In both aspects, our results point the importance of the additive component in accounting for international transport costs. In this respect, it contributes to the literature that challenges the dominant role of iceberg costs in international trade (even in the case of ice transport: see \citealp{bosker2018ice}). As underlined by \cite{alchian}, the relative price of two varieties of some good will depend on the level of additive trade costs. In this context, the relative demand for more expensive/higher quality product goods increases with trade costs (``shipping the good apples out''). This is consistent with the findings by \citet{hummels_skiba}, who estimate the elasticity of freight rates with respect to price to be well below unity. Also, their estimates imply that doubling freight costs increases average free alongside (fas) export prices by 80 to 141 percent, consistent with high quality goods being sold in markets with high freight costs. \citet{Lashkaripour-mimeo-2017} challenges this view. Highlighting that more expensive goods are systematically heavier (and hence more costly to transport\footnote{Yet, one can be concerned by the generality of this result. While the positive correlation between weight and price seems reasonable for goods from the second industrial revolution like cars, it is dubious in the case of ITC goods whose importance in international trade has been rising since the end of the twentieth century.}), he finds an average transport cost elasticity much closer to 1 using firm-level data on Colombian imports. This result is also confirmed on more aggregate US, sector-level trade data. The debate is not over yet, as recent empirical studies based on micro-level data conversely provide a strong empirical support to the role of additive costs (i.e., cost per unit exported) in international trade costs. Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher unit values for exports to more remote countries, supporting the importance of additive costs.\footnote{Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs. \citet{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive trade costs, in addition to the ad-valorem component. A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in ad-valorem costs, due to the alteration of relative prices in a heterogeneous-firms trade framework. This is confirmed by \cite{Irrazabal_2015}. While these results suggest that important welfare gains can be achieved by reducing additive trade costs, not much progress has been done in quantifying such gains. One potential reason is the lack of an empirical characterization of the additive component of trade costs (the one exception being \citet{Irrazabal_2015}). One contribution of the paper is to fill this gap.}

Closely related to our paper is the work by \citet{Irrazabal_2015}, which develop a structural framework for inferring relative additive trade costs from firm-level trade data. Implementing their methodology on Norwegian firm-level export data for the year 2004, they find that, for the median shipment, additive costs (in Norwegian crowns) amount to 6\% of the export price multiplied by the ad-valorem cost (expressed in Norwegian crowns). Our results share in common with \citet{Irrazabal_2015} the important role of the additive component of international trade costs. Yet, our paper complements their findings in two main respects. First, while our study covers a narrower set of trade costs focused on transport costs, their data and their empirical approach only allows the identification of the ratio of the additive cost to the export price multiplied by the ad-valorem cost ; by contrast, our estimation strategy enables us to uncover separately both values of the ad-valorem and the additive costs. From this, we can rebuilt the ratio in similar terms to them, thereby gaining in generality in this respect. While \citet{Irrazabal_2015} obtain that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs, our own results point (for 2004) to a value of 2.8 USD paid in additive transport costs for maritime shipping and 1.5 USD for air transport. This comparison of results suggests that between a third and a half of the additive trade costs are attributable to the transport sector (depending on the transport mode). \footnote{See Section \ref{sec:results_decomposition} for more details.} Second, we exploit exhaustive information about the imports flows of the US over a large time span from 1974 to 2013. In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time. In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed since 1974.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Kropf-Saure-JIE-2016}, which estimate the size and shape of per-shipment costs bases on Swiss data, and to \citet{Alessandria-et-al-AER-2010}, \citet{Hornok-et-al-RES-2015} and \citet{Hornok-et-al-JIE-2015}, which point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip

@@

The non-linearity of Equation (\ref{eq:equation0}) implies that it cannot be estimated using standard linear estimators. All estimates are thus performed using non-linear least squares.\footnote{The basis of the method is to approximate the model by a linear one and to refine the parameters by successive iterations. The intuitive criterion for convergence is that the sum of squares of residuals does not increase from one iteration to the next. See \cite{Woolridge-Book-2001} for more details.} Yet, the use of a nonlinear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time. Confronted to this arbitrage, we estimate international transport costs at the 3-digit level as our benchmark classification, even though data series are available at the 5-digit classification level ($k$). In other words, this amounts making the additional assumption, that all products $k$ in a 3-digit sector $s$ share the
same structure of costs.\footnote{One way to gauge the relevance of this assumption is to provide a decomposition variance exercise on the observed cif-fas price. As reported in Appendix \ref{app:decomp_variance}, the share of the observed variance that is accounted for by the between-sector ($s$) variance is roughly similar to the between-product ($k$) variance.}$^{,}$\footnote{One may object that we could preserve the estimation at the 5-digit level by running an OLS estimation on the equation taken in level, i.e. on the basis of Equation (\ref{eq:base_estimee}), specifying the error term additively. This would not solve the problem though, for three main complementary reasons. First, at the 5-digit level the number of fixed effects to include in the estimation would be more than 430,000 (with 216 countries and 2,029 products), making the estimation computationally extremely burdensome, even in OLS. Imposing Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}) to reduce the number of fixed effects would drive us back to the non-linearity issue for fixed effects. Second, making the error term (specified as following a normal law centered on 0) enter Equation (\ref{eq:base_estimee}) additively implies negative values for some of the estimated residuals $\widehat{\epsilon}_{ik}$, which is inconsistent with the constraint that $\frac{p_{ik}}{\widetilde{p}_{ik}}>1$. Third, estimating the equation in level does not eliminate the fact that it is non-linear by nature, as long as there are additive costs to estimate (see Equation (\ref{eq:base_estimee}) for $t_{ik} \neq 0$).}

@@

As mentioned in the Introduction, one contribution of the paper is to provide a careful estimation of international transport costs, that decomposes into both an additive and an ad-valorem component. In this respect, our paper relates to the recent literature about the importance of additive costs in accounting for international trade costs (\citealp{Irrazabal_2015}, among others). We contribute to this issue by adopting the following strategy. For each year and transport mode, we estimate two models: \textit{(a)} when additive costs are excluded, in which case transport costs are modeled as iceberg costs only (Equation (\ref{eq:model_nlI})), and \textit{(b)} when transport costs are decomposed in the two additive and ad-valorem dimensions (Equation (\ref{eq:model_IetA})). From this, we compare the fitting properties and the explanatory power of both models to evaluate the importance of modelling the additive component.

Under Model  \textit{(a)}, the estimated equation simplifies as:\footnote{One may object that a comprehensive study of the structure of transport costs should also include the third model with only additive costs. This has driven us to estimate this model as well, in which case the estimated equation is written according to: $\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}\right) + \epsilon^{add}_{ik}$. The main result that emerges is that the model with additive costs only is dominated (in terms of quality of fit properties) by the model with multiplicative costs only (Equation (\ref{eq:model_nlI})), which is itself dominated by the complete model (Equation (\ref{eq:model_IetA})), anticipating on further results. These results are reported in the Online Appendix, available on the authors' webpages.}

@@

@@

After estimating Equation (\ref{eq:model_IetA}), we can re-built a measure of each component, $\widehat{\tau}^{adv}_{is} = \widehat{\tau_{i}} \times \widehat{\tau}_{s(k)}$ for the ad-valorem cost and $\widehat{t}_{is} = \widehat{t}_{i} + \widehat{t}_{s(k)}$ for the additive cost, that are country-sector specific, by year and transport mode. When assuming iceberg costs only (Equation (\ref{eq:model_nlI})), we proceed similarly to get $\widehat{\tau}^{ice}_{is} = \widehat{\tau}_{i} \times \widehat{\tau}_{s}$. In this case, notice that the equation could be estimated relying on a linear form. To preserve comparability of the results, we keep the same non-linear estimation method in both cases though. Similarly as \citet{Irrazabal_2015}, we take the average over the sector-country dimension, using the values of each trade flow ($is$-specific) over total yearly trade as a weighting scheme. We thus recover a ``synthetic estimate'' of each type of transport cost: $\widehat{\tau}^{ice}$ for model \emph{(a)}, $\widehat{\tau}^{adv}$ and $\widehat{t}$ for model \textit{b)}, for each year and transportation mode. These results are reported in Section \ref{sec:results_decomposition}.

@@

Our ambition to re-assess the underlying sources of the trend patterns of international transport costs requires a careful modeling of both the additive and the multiplicative components. The first step of the analysis then consists in providing a quantification of the magnitude of transport costs over time (by transport mode), distinguishing whether the additive component $t_{ik}$ is excluded or included in the estimated equation (Equation (\ref{eq:model_IetA}) or (\ref{eq:model_nlI})). This allows us to provide estimates for the size of both the ad-valorem and the additive components of transport costs, and assess whether additive costs represent a sizeable component of the latter. In this respect, these results \textit{per se} also constitute our first original contribution to the literature. In Appendix \ref{app:diagnostic_test}, we provide diagnostic tests about the goodness-of-fit measures, that confirm the empirical relevance of adding the additive component in the estimation.\smallskip

Table \ref{tab:summary_results} reports a summary of our results. It displays the mean and median values of each type of transport costs (either ad-valorem estimated alone or estimated along with additive costs), as well as the associated standard deviation, averaged over the period 1974-2013, for estimation driven both at the 3- and 4-digit sectorial level for different specifications and data. The first panel reports the results for a specification based on ad-valorem costs estimated alone (Equation (\ref{eq:model_nlI})), while the second panel presents estimates for a specification involving both ad-valorem and multiplicative components (Equation (\ref{eq:model_IetA})). Finally, the third panel reports the same set of descriptive statistics, but for the actual cif/fas ratio in our data.\footnote{In Appendix \ref{app:more_results}, we report similar results for a sample of years, for both transport mode, at the 3 and 4-digit classification level. Results for all years (available at the 3-digit level) are reported in the Online Appendix, available on the authors' webpages.}$^{,}$\footnote{We present the estimation results for Air at the 3-digit level removing the year 1989, as the results reveal the presence of strong outliers that bias the estimates of transport costs upwards this particular year. Overall results (over the whole period) are not substantially affected if this year is included though. This can be uncovered from the results detailed year by year in the Online Appendix, available on the authors' webpages.}

>>>>

\begin{abstract}
This paper investigates the underlying sources behind the downward trend of international transport costs. In the same spirit as \cite{hummels2007}, we identify the respective roles of the reduction in ``pure'' transport costs and changes in trade composition effects. This drives us to provide a careful modeling of international transport costs, that allows for both ad-valorem (multiplicative) and per-unit (additive) components, exploiting information contained in the US imports flows database over 1974-2013 distinguished by transport mode, air and vessel. In contrast to \cite{hummels2007}, trade composition effects do not matter much and when they do, they tend to amplify (rather than reduce) the decrease in pure transport costs. Importantly, this difference of results can be attributed to the new way of modeling the additive component of transport costs we offer. Further, in quantitative terms, we find that the additive component of transport costs is sizeable, representing between 40 and 50\% of overall transport costs. In both aspects, our results point the importance of the additive component in accounting for international transport costs.
\end{abstract}

@@

This focus is shared with part of the international trade literature that studies the patterns of transport costs over time, such as \cite{hummels2007} and \cite{Behar_Venables} (among others). Many argue that transport costs have substantially decreased with technological advance in transportation (\citealp{levinson2016box}), infrastructure development and new communication technologies (\citealp{Lafourcade_Thisse}). This view does not make consensus though. Distance as an incompressible cost is still a very important determinant of trade flows (\textbf{citer?}); further, progress in transportation techniques should be compared with technical progress in production to get a clear picture of the transport costs burden. The constant-dollar cost of transport has declined, but so has the constant-dollar cost of production. As well, the quality of transport services has increased. In accordance with this view, \cite{hummels2007} points out that transport costs in proportion of export prices have declined less quickly than expected.\smallskip

According to \cite{hummels2007}, this lower-than-expected decline of transport costs is partly explained by trade composition effects (i.e., changes over time in the baskets of imported goods and/or the origin countries). Based on US imports over 1974-2004, he finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs (i.e., \textit{ceteris paribus} for given goods, mode, and trade routes) because trade composition effects have partly offset the ``pure'' transport costs decrease, thereby attenuating the overall transport costs downward trend.\footnote{Throughout the paper, we refer to ``pure'' or ``per se'' transport costs changes as the changes over time in the transport costs for a given product/country partner, ie excluding trade composition effects.} In this paper, we investigate the robustness of this result. As a major contribution, and in contrast to \cite{hummels2007}, we find that trade composition effects play a minor role in accounting for the trend patterns of transport costs. Rather, the evolution of transport costs ``per se'' constitutes the main driving force behind the overall transport costs decrease, in particular for air transport. In maritime shipment, trade composition effects matter more, but they amplify (rather than offset) the reduction in pure transport costs, suggesting that pure transport costs declined even less than argued by \cite{hummels2007}.\medskip

Proving this point drives us to provide a careful modelling of transport costs. Following \citet{samuelson1954}, standard models of international trade have usually modeled trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis).\footnote{Strictly speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded. As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as it is commonly done in the literature.} However, a recent strand of the literature points out to the existence of an additive component, that is, a cost per unit traded (see, among others, \cite{Irrazabal_2015} or \citealp{martin2012}). Based on US sectoral data, our own estimates also bring out the importance of additive costs in international transport costs. This is obtained by providing an empirical decomposition of the structure of transport costs over time, in which we explicitly distinguish between ad-valorem and additive costs. To do so, we exploit information contained in the US imports flows over 1974-2013, from which we get a measure of international transport costs as the difference between the import and the export prices, on a year-product-partner country basis. This database is the same as \cite{hummels2007}, extended over more recent years (2005-2013). \smallskip

@@

Second, we find that the main source of the downward trend in overall transport costs (that decreased by roughly 50\% in air shipping and 60\% in vessel shipping over 1974-2013) does come from the reduction of transport cost ``per se''. Trade composition effects only have a limited influence, and when they matter (in maritime transport), they amplify the reduction of pure transport costs, suggesting that pure transport costs declined less than observed costs. This stands in contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004). Importantly, we show that the difference of results can be attributed to the new way of modeling of the additive component we offer. Allowing the share of additive costs to vary across time, products and country partners (rather than being constant as in \citealp{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period. Our result thus deepens the paradox of the less-than-expected decline of transport costs. \textbf{j'aime bien parler de ce paradoxe mais la seule source qu'on cite c'est Hummels 2007, d'autres papiers?}\smallskip

In both aspects, our results point the importance of the additive component in accounting for international transport costs. In this respect, it contributes to the literature that challenges the dominant role of iceberg costs in international trade (even in the case of ice transport, see \citealp{bosker2018ice}). As underlined by \cite{alchian}, the relative price of two varieties of some good will depend on the level of additive trade costs. In this context, the relative demand for more expensive/higher quality product goods increases with trade costs (``shipping the good apples out''). This is consistent with the findings by \citet{hummels_skiba}, who estimate the elasticity of freight rates with respect to price to be well below unity. Also, their estimates imply that doubling freight costs increases average free alongside (fas) export prices by 80 to 141 percent, consistent with high quality goods being sold in markets with high freight costs. \citet{Lashkaripour-mimeo-2017} challenges this view. Highlighting that more expensive goods are systematically heavier (and hence more costly to transport\footnote{Yet, one can be concerned by the generality of this result. While the positive correlation between weight and price seems reasonable for goods from the second industrial revolution like cars, it is dubious in the case of ITC goods whose importance in international trade has been rising since the end of the twentieth century.}), he finds an average transport cost elasticity much closer to 1 using firm-level data on Colombian imports. This result is also confirmed on more aggregate US, sector-level trade data. The debate is not over yet, as recent empirical studies based on micro-level data conversely provide a strong empirical support to the role of additive costs (i.e., cost per unit exported) in international trade costs. Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher unit values for exports to more remote countries, supporting the importance of additive costs.\footnote{Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs. \citet{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive trade costs, in addition to the ad-valorem component. A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in ad-valorem costs, due to the alteration of relative prices in a heterogeneous-firms trade framework. This is confirmed by \cite{Irrazabal_2015}. While these results suggest that important welfare gains can be achieved by reducing additive trade costs, not much progress has been done in quantifying such gains. One potential reason is the lack of an empirical characterization of the additive component of trade costs (the one exception being \citet{Irrazabal_2015}). One contribution of the paper is to fill this gap.}

Closely related to our paper is the work by \citet{Irrazabal_2015}, which develop a structural framework for inferring relative additive trade costs from firm-level trade data. Implementing their methodology on Norwegian firm-level export data for the year 2004, they find that, for the median shipment, additive costs (in Norwegian crowns) amount to 6\% of the export price multiplied by the ad-valorem cost (expressed in Norwegian crowns). Our results share in common with \citet{Irrazabal_2015} the important role of the additive component of international trade costs. Yet, our paper complements their findings in two main respects. First, while our study covers a narrower set of trade costs focused on transport costs, their data and their empirical approach only allows the identification of the ratio of the additive cost to the export price multiplied by the ad-valorem cost ; by contrast, our estimation strategy enables us to uncover separately both values of the ad-valorem and the additive costs. From this, we can rebuilt the ratio in similar terms to them, thereby gaining in generality in this respect. While \citet{Irrazabal_2015} obtain that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs, our own results point (for 2004) to a value paid in additive transport costs of 2.8 USD for maritime shipping and 1.5 USD for air transport. This comparison of results suggests that between a third and a half of the additive trade costs are attributable to the transport sector (depending on the transport mode). Second, we exploit exhaustive information about the imports flows of the US over a large time span from 1974 to 2013. In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time. In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed since 1974.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Kropf-Saure-JIE-2016}, which estimate the size and shape of per-shipment costs bases on Swiss data, and to \citet{Alessandria-et-al-AER-2010}, \citet{Hornok-et-al-RES-2015} and \citet{Hornok-et-al-JIE-2015}, which point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip

@@

The non-linearity of Equation (\ref{eq:equation0}) implies that it cannot be estimated using standard linear estimators. All estimates are thus performed using non-linear least squares.\footnote{The basis of the method is to approximate the model by a linear one and to refine the parameters by successive iterations. The intuitive criterion for convergence is that the sum of squares of residuals does not increase from one iteration to the next. See \cite{Woolridge-Book-2001} for more details.} Yet, the use of a nonlinear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time. Confronted to this arbitrage, we estimate international transport costs at the 3-digit level as our benchmark classification, even though data series are available at the 5-digit classification level ($k$). In other words, this amounts making the additional assumption, that all 5-digit products $k$ in a 3-digit sector $s$ share the
same structure of costs.\footnote{One way to gauge the relevance of this assumption is to provide a decomposition variance exercise on the observed cif-fas price. As reported in Appendix \ref{app:decomp_variance}, the share of the observed variance that is accounted for by the between-sector ($s$) variance is roughly similar to the between-product ($k$) variance.}$^{,}$\footnote{One may object that we could preserve the estimation at the 5-digit level by running an OLS estimation on the equation taken in level, i.e. on the basis of Equation (\ref{eq:base_estimee}), specifying the error term additively. This would not solve the problem though, for three main complementary reasons. First, at the 5-digit level the number of fixed effects to include in the estimation would be more than 430,000 (with 216 countries and 2,029 products), making the estimation computationally extremely burdensome, even in OLS. Imposing Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}) to reduce the number of fixed effects would drive us back to the non-linearity issue for fixed effects. Second, making the error term (specified as following a normal law centered on 0) enter Equation (\ref{eq:base_estimee}) additively implies negative values for some of the estimated residuals $\widehat{\epsilon}_{ik}$, which is inconsistent with the constraint that $\frac{p_{ik}}{\widetilde{p}_{ik}}>1$. Third, estimating the equation in level does not eliminate the fact that it is non-linear by nature, as long as there are additive costs to estimate (see Equation (\ref{eq:base_estimee}) for $t_{ik} \neq 0$).}

@@

As mentioned in the Introduction, one contribution of the paper is to provide a careful estimation of international transport costs, that decomposes into both an additive and an ad-valorem component. In this respect, our paper relates to the recent literature about the importance of additive costs in accounting for international trade costs (\citealp{Irrazabal_2015}, among others). We contribute to this issue by adopting the following strategy. For each year and transport mode, we estimate two models. In Model (a), additive costs are excluded, transport costs being modeled as iceberg costs only (Equation (\ref{eq:model_nlI})). In this case, the estimated equation simplifies as:

@@

Under Model (b), transport costs are decomposed in the two additive and ad-valorem dimensions (Equation (\ref{eq:model_IetA})). From this, we compare the fitting properties and the explanatory power of both models to evaluate the importance of modelling the additive component.\footnote{One may object that a comprehensive study of the structure of transport costs should also include the third model with only additive costs. This has driven us to estimate this model as well, in which case the estimated equation is written according to: $\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}\right) + \epsilon^{add}_{ik}$. The main result that emerges is that the model with additive costs only is dominated (in terms of quality of fit properties) by the model with multiplicative costs only (Equation (\ref{eq:model_nlI})), which is itself dominated by the complete model (Equation (\ref{eq:model_IetA})), anticipating on further results. These results are reported in the Online Appendix, available on the authors' webpages.}

@@

After estimating Equation (\ref{eq:model_IetA}), we can re-built a measure of each component, $\widehat{\tau}^{adv}_{is} = \widehat{\tau_{i}} \times \widehat{\tau}_{s(k)}$ for the ad-valorem cost and $\widehat{t}_{is} = \widehat{t}_{i} + \widehat{t}_{s(k)}$ for the additive cost, that are country-sector specific, by year and transport mode. When assuming iceberg costs only (Equation (\ref{eq:model_nlI})), we proceed similarly to get $\widehat{\tau}^{ice}_{is} = \widehat{\tau}_{i} \times \widehat{\tau}_{s}$. In this case, notice that the equation could be estimated relying on a linear form. To preserve comparability of the results, we keep the same non-linear estimation method in both cases though. Similarly as \citet{Irrazabal_2015}, we take the average over the sector-country dimension, using the values of each trade flow ($is$-specific) over total yearly trade as a weighting scheme. We thus recover a ``synthetic estimate'' of each type of transport cost: $\widehat{\tau}^{ice}$ for Model (a), $\widehat{\tau}^{adv}$ and $\widehat{t}$ for Model (b), for each year and transportation mode. These results are reported in Section \ref{sec:results_decomposition}.

@@

Our ambition to re-assess the underlying sources of the trend patterns of international transport costs requires a careful modeling of both the additive and the multiplicative components. The first step of the analysis then consists in providing a quantification of the magnitude of transport costs over time (by transport mode), distinguishing whether the additive component $t_{ik}$ is excluded or included in the estimated equation (Equation (\ref{eq:model_IetA}) or (\ref{eq:model_nlI})). This allows us to provide estimates for the size of both the ad-valorem and the additive components of transport costs, and assess whether additive costs represent a sizeable component of the latter. These results constitute our first original contribution to the literature. In Appendix \ref{app:diagnostic_test}, we provide diagnostic tests about the goodness-of-fit measures, that confirm the empirical relevance of adding the additive component in the estimation.\smallskip

Table \ref{tab:summary_results} reports a summary of our results. It displays the mean and median values of each type of transport costs (either ad-valorem estimated alone or estimated along with additive costs), as well as the associated standard deviation, averaged over the period 1974-2013, for estimation driven both at the 3- and 4-digit sectorial level for different specifications and data. The top panel reports the results for a specification based on ad-valorem costs estimated alone (Model (a), based on Equation (\ref{eq:model_nlI})), while the intermediate panel presents estimates for a specification involving both ad-valorem and multiplicative components (Model (b), based on Equation (\ref{eq:model_IetA})). Finally, the bottom panel reports the same set of descriptive statistics, but for the actual cif/fas ratio in our data.\footnote{In Appendix \ref{app:more_results}, we report similar results for a sample of years, for both transport mode, at the 3- and 4-digit classification level. Results for all years (available at the 3-digit level) are reported in the Online Appendix, available on the authors' webpages.}$^{,}$\footnote{We present the estimation results for Air at the 3-digit level removing the year 1989, as the results reveal the presence of strong outliers that bias the estimates of transport costs upwards this particular year. Overall results (over the whole period) are not substantially affected if this year is included though. This can be uncovered from the results detailed year by year in the Online Appendix, available on the authors' webpages.}

\begin{table}[htbp]
  \centering
  \footnotesize{
  \caption{Transport costs estimates: Summary \label{tab:summary_results}}
  \begin{center}
    \begin{tabular}{l|cc|cc}
      \hline \hline
    \multicolumn{5}{c}{Mean value over 1974-2013}   \\
    \# digit & \multicolumn{2}{c}{3 digits} & \multicolumn{2}{c}{4 digits} \\ \hline
    Mode  & Vessel & Air ($^{\ast \ast}$) & Vessel & Air \\ \hline
    \multicolumn{5}{l}{\textbf{Model (a) - With only Ad-Valorem Transport Costs} ($\widehat{\tau}^{ice} -1$, in \%)}  \\ \hline
    Mean  & 5.8 & 5.1 & 6.0 & 4.9 \\
    Median & 5.1 & 4.2 & 5.2 & 3.7 \\ \hline
    %Std   & 0.032 & 0.042 & 0.036 & 0.045 \\
    %Min. value & 1.003 & 1.001 & 1.003 & 1.000 \\
    %Max. value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
    \multicolumn{5}{l}{\textbf{Model (b) - With Additive \& Ad-Valorem Transport Costs} } \\ \hline
   \textit{Ad-valorem term ($\widehat{\tau}^{adv}-1$, in \%)} & & & & \\ \hline
    Mean  & 3.2 & 2.5 & 3.3 & 2.4 \\
    Median & 2.8 & 1.8 & 2.8 & 1.6 \\ \hline
    %Std   & 0.023 & 0.023 & 0.025 & 0.026 \\
    %Min. value & 1.001 & 1.000 & 1.000 & 1.000 \\
    %Max. value & 1.227 & 1.474 & 1.264 & 1.537 \\ \hline
    \textit{Additive term }& & & &   \\ \hline
    \textit{In \% of the export price ($\widehat{t}/\widetilde{p}$, in \%)} &&&& \\ \hline
        Mean  & 2.9 & 1.8 & 2.8 & 1.9 \\
    Median & 1.9 & 0.7 & 1.7 & 0.8 \\ \hline
   \textit{In USD per kg traded ($\widehat{t}$)}&&&& \\ \hline
    Mean & 0.09	&1.12	&0.08&	2.16 \\
    Median & 0.07	&1.04	&0.06&	1.06 \\ \hline
    %Std   & 0.041 & 0.034 & 0.039 & 0.034 \\
    %Min. value & 0.000 & 0.000 & 0.000 & 0.000 \\
    %Max. value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
    \multicolumn{5}{l}{\textbf{Data}  } \\ \hline
    Transport costs ($p/\widetilde{p} -1$, in \%) & & & & \\ \hline
    Mean & 5.3 & 5.0& 5.6&3.9 \\
    Median & 4.3 & 2.0 & 4.4& 1.9 \\ \hline
    Export price ($\widetilde{p}$), in USD & & & & \\
    Mean & 16.0 &	6488.4	&9.6	&6643.6 \\
    Median & 4.1	& 142.5	& 4.1	& 142.2 \\ \hline
    \# obs. & 29279 & 28207 & 29317 & 27680 \\
    \# origin country & 188 & 191 & 188 & 189 \\
    \# products & 230 & 211 & 666 & 567 \\  \hline \hline
  \end{tabular}
    \end{center}}
\parbox[l]{10cm}{\tiny{Notes: Statistics are obtained weighting each observation by its value relative to total trade flows. The additive term is expressed in fraction of fas price. For the  4-digit classification, statistics for observed data have been calculated for the same set of years as used for estimation, i.e. 1974, 1981, 1989, 2001, 2009, 2013. ($^{\ast \ast}$): 1989 omitted in 3-digit estimation for air.}}
\end{table}%

Table \ref{tab:summary_results} calls for two types of comments. The first type is of statistical order. As displayed in Table \ref{tab:summary_results}, estimated mean values are systematically higher than medians, by around one percentage point. This result is not surprising recalling that our key variable is by nature bounded to 0; therefore, outliers can only be positive, pushing the mean upward compared to the median. One can also note that estimated total costs are systematically higher than observed costs, by 0.5 to 1 percentage point. This is first due to the fact that observed data are \textit{de facto} trade-weighed (i.e., by the weight of each good in total trade) while our estimates are based on simple arithmetic means within sector and country. Furthermore, Table \ref{tab:summary_results} reports values expressed in percentages (for the observed values, $\frac{p_{ik}}{\widetilde{p}_{ik}} -1$, in \%), whereas all our estimates are based on log-linearization. Therefore, what matters on the statistical ground is that the mean of our predicted values (based on logarithms) must be equal to the mean of the log-linearized data ($\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)$), which does not necessarily imply equality between values. Descriptive statistics showing that that both observed and predicted values do match in logarithms are available upon request to the authors.\medskip
